Controlling synchronization in coupled area-preserving maps by stickiness

### 2018-09-26 Syncarea_control/paper.tex
paper and ref updated
paper updated

\noindent Here the subscript $n$ denotes the discrete time and $K$ is the 
nonlinearity parameter. These equations typically describe the evolution of 
two canonical variables $P$ and $Q$ which correspond to the momentum and 
co-ordinate in the Poincar\'{e} section of a freely moving rotator with 
interleaved periodic kicks. This system represents the behavior of a variety 
of systems such as charged particle confinement in  mirror magnetic traps, 
particle dynamics in accelerator, comet dynamics in solar systems 
etc.~\cite{Chirikov1960,Izraelev1980,Chirikov1989,Chirikov2008}
Two-dimensional  phase space plots  of the standard map for the parameter 
value $K = 6$ using $25$ initial conditions  are shown in 
Figs.~\ref{fig:Standard_map}.

@@

We now analyze synchronization of the coupled system using the master stability function~\cite{Pecora1998}. The numerical procedures to control synchronization times will be discussed thereafter.

@@

\begin{align}
\frac{d\overbar{X}_d}{dt} & = \overbar{F}(\overbar{X}_d) \nonumber  \\
\frac{d\overbar{X_r}}{dt} &= \overbar{F}(\overbar{X}_r) + \alpha 
E(\overbar{X}_d-\overbar{X}_r).
\end{align}

@@

\begin{align}
\overbar{X}^d_{n+1} &= \overbar{F}(\overbar{X}^d_{n}) \nonumber  \\
\overbar{X}^r_{n+1} &= \overbar{F}(\overbar{X}^r_{n}) + \alpha 
E(\overbar{X}^d_{n}-\overbar{X}^r_n).
\end{align}

Therefore, in the case of a general unidirectional coupling of two standard maps, we get
\begin{align}
P^d_{n+1} &= P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
Q^d_{n+1} &= P^d_{n+1} + Q^d_n \nonumber\\
P^r_{n+1} &= P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
Q^r_{n+1} &= P^r_{n+1} + Q^r_n.
\label{equ:coupled}
\end{align}

@@

To find the stability of the synchronous state, we first express  
Eq.(\ref{equ:coupled}) in terms of  $P^\perp = P^\textnormal{d} - 
P^\textnormal{r}$ and $Q^\perp = Q^\textnormal{d} -Q^\textnormal{r}$, as 
follows
\begin{eqnarray}
\label{equ:trans}
P^\perp_{n+1} &= (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^d_n) - 
\frac{K}{2\pi} \sin(2\pi Q^r_n) \nonumber \\ 
Q^\perp_{n+1} &= (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi 
Q^d_n) \nonumber \\ &- \frac{K}{2\pi} \sin(2\pi Q^r_n).
\end{eqnarray}

We now write the variational equation for Eq.(\ref{equ:trans}) by linearizing 
about $(P^\textnormal{d}_n,Q^\textnormal{d}_n)$
\begin{equation}
\label{equ:variational}
\left[ \begin{array}{c} \delta P^\perp_{n+1} \\ \delta Q^\perp_{n+1} 
\end{array} \right] = \mathcal{M}(\alpha)\left[ \begin{array}{c}\delta 
P^\perp_{n} \\ \delta Q^\perp_{n} \end{array} \right],
\end{equation}
where the matrix $\mathcal{M}(\alpha)$ is given by
\begin{equation} 
\label{equ:M}
\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^d_n) \\ 1 - 
\alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix}.
\end{equation}

@@

\begin{equation} 
\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^d_n) \\ 0 & 1 + K\cos(2\pi 
Q^d_n) \end{bmatrix},
\end{equation}

@@

\begin{equation}
\label{equ:LE}
\lambda   = \lim_{n \rightarrow \infty} 
\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^d_n)|.
\end{equation}

@@

\begin{minipage}[t]{0.45\textwidth}
	%\raggedleft
	\centering
	\[ J = \left( \begin{array}{cc}
	1 & 1 + (K-\epsilon)\cos(2\pi Q_n)   \\
	1 & (K-\epsilon)\cos(2\pi Q_n) \end{array} \right)\]
\end{minipage}

@@

\section{Conclusions}
\label{sec:conclusions}
To summarize, we have developed numerical procedures to control 
synchronization in two identical standard maps coupled under the 
drive-response configuration. The procedure is based on the fact that 
synchronization in the system typically occurs in the sticky neighborhoods of the regular islands in 
the phase space. We have shown that a delay in synchronization can be obtained 
by kicking the trajectories slightly away from the domain temporarily. The 
efficiency of the procedure depends upon the number of steps employed in 
control. By applying the procedure to the system on merely $1\%$ of the phase 
space area (at $K = 6$), we have successfully increased synchronization times. 
The maximum number of control steps that are used is $0.01\%$ of the total of 
number of iterations used throughout the computation.  Furthermore, the 
procedure gives variety of delays. The limitations of this procedure are -- (a) 
the deflection may insert the drive trajectory in the regular islands which 
would lead to undesirable faster synchronization, and (b) a predetermined 
synchronization time in not obtainable. Furthermore, we have demonstrated a 
simple way to fasten synchronization by the parameter perturbation technique 
which rapidly drives the chaotic trajectories in the sticky neighborhood of 
regular islands. The distribution of degree of controlled synchronization 
times indicates the efficiency of the procedure. In both the numerical 
procedures, we have ignored the complex hierarchical structures which lead to 
sticky behavior of the chaotic orbits in the phase space. A possible future 
direction is to investigate the role of stickiness in measure synchronization 
in coupled Hamiltonian systems.

>>>>

\noindent Here the subscript $n$ denotes the discrete time and $K$ is the 
nonlinearity parameter. These equations typically describe the evolution of 
two canonical variables $P$ and $Q$ which correspond to the momentum and 
co-ordinate in the Poincar\'{e} section of a freely moving rotator with 
interleaved periodic kicks. This system represents the behavior of a variety 
of systems such as charged particle confinement in  mirror magnetic traps, 
particle dynamics in accelerator, comet dynamics in solar systems 
etc.~\cite{Chirikov1960,Izraelev1980,Chirikov1989,Chirikov2008}
Two-dimensional  phase space plots  of the standard map for the parameter 
value $K = 6$ using $25$ initial conditions  are shown in 
Fig.~\ref{fig:Standard_map}.

@@

We now analyze synchronization of the coupled system using the master stability function~\cite{Pecora1998}. Numerical procedures to control synchronization times will be discussed thereafter.

@@

\begin{align}
\frac{d\overbar{X}_\textnormal{d}}{dt} & = \overbar{F}(\overbar{X}_\textnormal{d}) \nonumber  \\
\frac{d\overbar{X_\textnormal{r}}}{dt} &= \overbar{F}(\overbar{X}_\textnormal{r}) + \alpha 
E(\overbar{X}_\textnormal{d}-\overbar{X}_\textnormal{r}).
\end{align}

@@

\begin{align}
\overbar{X}^\textnormal{d}_{n+1} &= \overbar{F}(\overbar{X}^\textnormal{d}_{n}) \nonumber  \\
\overbar{X}^\textnormal{r}_{n+1} &= \overbar{F}(\overbar{X}^\textnormal{r}_{n}) + \alpha 
E(\overbar{X}^\textnormal{d}_{n}-\overbar{X}^\textnormal{r}_n).
\end{align}

Therefore, in the case of a general unidirectional coupling of two standard maps, we get
\begin{align}
P^\textnormal{d}_{n+1} &= P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi Q^\textnormal{d}_n) \nonumber\\
Q^\textnormal{d}_{n+1} &= P^\textnormal{d}_{n+1} + Q^\textnormal{d}_n \nonumber\\
P^\textnormal{r}_{n+1} &= P^\textnormal{r}_n + \frac{K}{2\pi}\sin (2\pi Q^\textnormal{r}_n) + \alpha(P^\textnormal{d}_n-P^\textnormal{r}_n) \nonumber \\
Q^r_{n+1} &= P^r_{n+1} + Q^r_n.
\label{equ:coupled}
\end{align}

@@

To find the stability of the synchronous state, we first express  
Eq.(\ref{equ:coupled}) in terms of  $P^\perp = P^\textnormal{d} - 
P^\textnormal{r}$ and $Q^\perp = Q^\textnormal{d} -Q^\textnormal{r}$, as 
follows
\begin{eqnarray}
\label{equ:trans}
P^\perp_{n+1} &= (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^\textnormal{d}_n) - 
\frac{K}{2\pi} \sin(2\pi Q^\textnormal{r}_n) \nonumber \\ 
Q^\perp_{n+1} &= (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi 
Q^\textnormal{d}_n) \nonumber \\ &- \frac{K}{2\pi} \sin(2\pi Q^\textnormal{r}_n).
\end{eqnarray}

We now write the variational equation for Eq.(\ref{equ:trans}) by linearizing 
about $(P^\textnormal{d}_n,Q^\textnormal{d}_n)$
\begin{equation}
\label{equ:variational}
\left[ \begin{array}{c} \delta P^\perp_{n+1} \\ \delta Q^\perp_{n+1} 
\end{array} \right] = \mathcal{M}(\alpha)\left[ \begin{array}{c}\delta 
P^\perp_{n} \\ \delta Q^\perp_{n} \end{array} \right],
\end{equation}
where the matrix $\mathcal{M}(\alpha)$ is given by
\begin{equation} 
\label{equ:M}
\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^\textnormal{d}_n) \\ 1 - 
\alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix}.
\end{equation}

@@

\begin{equation} 
\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^\textnormal{d}_n) \\ 0 & 1 + K\cos(2\pi 
Q^d_n) \end{bmatrix},
\end{equation}

@@

\begin{equation}
\label{equ:LE}
\lambda   = \lim_{n \rightarrow \infty} 
\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^\textnormal{d}_n)|.
\end{equation}

@@

\begin{minipage}[t]{0.45\textwidth}
	%\raggedleft
	\centering
	\[ J = \left( \begin{array}{cc}
	1 & 1 + (K-\epsilon)\cos(2\pi Q_n)   \\
	1 & (K-\epsilon)\cos(2\pi Q_n) \end{array} \right),\]
\end{minipage}

@@

\section{Conclusions}
\label{sec:conclusions}
To summarize, we have developed numerical procedures to control 
synchronization in two identical standard maps coupled under the 
drive-response configuration. The procedure is based on the fact that 
synchronization in the system typically occurs in the sticky neighborhoods of the regular islands in 
the phase space. We have shown that a delay in synchronization can be obtained 
by kicking the trajectories slightly away from the domain temporarily. The 
efficiency of the procedure depends upon the number of steps employed in 
control. By applying the procedure to the system on almost $16\%$ of the phase 
space area (at $K = 6$), we have successfully increased synchronization times. 
The maximum number of control steps that are used is $0.01\%$ of the total of 
number of iterations used throughout the computation.  Furthermore, the 
procedure gives variety of delays. The limitations of this procedure are -- (a) 
the deflection may insert the drive trajectory in the regular islands which 
would lead to undesirable faster synchronization, and (b) a predetermined 
synchronization time in not obtainable. Furthermore, we have demonstrated a 
simple way to fasten synchronization by the parameter perturbation technique 
which rapidly drives the chaotic trajectories in the sticky neighborhood of 
regular islands. The distribution of degree of controlled synchronization 
times indicates the efficiency of the procedure. In both the numerical 
procedures, we have ignored the complex hierarchical structures which lead to 
sticky behavior of the chaotic orbits in the phase space. A possible future 
direction is to investigate the role of stickiness in measure synchronization 
in coupled Hamiltonian systems.
